---
title: "Hosting pages on Github using Rmarkdown and Hugo"
author: "Guanghua Wang"
date: 2017-02-18T20:50:00
categories: ["Tech"]
tags: ["R","R Markdown","R Blogdown", "Hugo", "github"]
summary: "My personal website is created by R package `blogdown` and `Hugo` and is hosted on Github. This page summarise my steps to create this site."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
My personal website is created by R package `blogdown` and `Hugo`[^1] and is hosted on Github. This page summarise my steps to setup the site. 

# Creating github repositories
Two respositories are needed for the site.

1. I create `blog_hugo` to host Hugo’s content)

2. I create `<my-user-name>.github.io` to host the public folder, which contains the static website.

# Installing Blogdown and Hugo
## Install [blogdown](https://github.com/rstudio/blogdown).
It is under heavy development when I write this blog.
```{r install blogdown,eval=FALSE}
# check devetolls
if (!("devtools" %in% installed.packages())) {
  install.packages("devtools")
}
devtools::install_github('rstudio/blogdown')
```

## Install [Hugo](https://gohugo.io/)
As I use Deepin Linux, a debain-based distro, I install `Hugo` via `apt` package manager. For other installment choices, read the [Hugo installation instructions](https://gohugo.io/overview/installing/).
```
sudo apt install hugo
```
However, the Hugo in the repository is updated and it does not work for most templates. I manually download its deb package from [the release page](https://github.com/spf13/hugo/releases) and install it.

# Creating the Site
## Creating a Rproject of my new site
1. Clone `blog_hugo` to my computer
```{bash, eval = F}
cd <parent-folder-of-blog_hugo>
git clone <my-blog-hugo-repository> && cd blog_hugo
```
2. Create a Rproject based on the existing folder `blog_hugo`. It is a R project with the git system by default.

## Creating a new site
1. Pick a Hugo theme from [http://themes.gohugo.io](http://themes.gohugo.io). I use **Academic** for my site.
2. Creating a new site
```{r, eval=FALSE}
blogdown::new_site(dir = ".", theme = "gcushen/hugo-academic")
```

# Hosting the Site on Github
Use the [submodule](https://gist.github.com/gitaarik/8735255) to sync my local sites(the public folder) with `<my-user-name>.github.io` 
```{bash, eval=F}
git submodule add -b master git@github.com:<<my-user-name>/<my-user-name>.github.io.git public
```
As `SSH` is used to sync with Github, I need to create my own private key pair set in advance.
```
cd ~/.ssh && ssh-keygen
cat id_rsa.pub
```
Then, we copy the key to our github account via the website. More information can be found [here](http://stackoverflow.com/questions/2643502/git-permission-denied-publickey)

# Creating Scripts
Add a `serve_site.R` script in `blog_hugo` folder to generate static sites in case Rstudio is not running.
```
#!/usr/bin/Rscript
blogdown::serve_site()
```

Add a `deploy.sh` script in `blog_hugo` foler to sync the site (and make it executable: `chmod +x deploy.sh`)
```
# Go To Public folder
cd public
# Add changes to git.
git add -A

# Commit changes.
msg="rebuilding site `date`"
if [ $# -eq 1 ]
  then msg="$1"
fi
git commit -m "$msg"

# Push source and build repos.
git push origin master

# Come Back
cd ..
```

# The Last Step
Modify the theme as we like and enjoy the new site.

# Reference
1. [Hosting on Github Pages](https://gohugo.io/tutorials/github-pages-blog/)

2. [blogdown](https://github.com/rstudio/blogdown)

3. [hugo-academic](https://github.com/gcushen/hugo-academic)

[^1]: `Hugo` can be used to hosting other types of pages on Github. More information can be found on [its website](https://gohugo.io/tutorials/github-pages-blog/)
